---
title       : Data Structures & Subsetting
subtitle    : Based on material by Hadley Wickham
author      : Sunanda Garg
job         : 
logo        : r_image.png
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [quiz, bootstrap]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---

## Data Structures[^1]

|    | Homogeneous   | Heterogeneous |
|----|---------------|---------------|
| 1d | Atomic vector | List          |
| 2d | Matrix        | Data frame    |
| nd | Array         |               |
 
#### Source: [Advanced R Programming](http://adv-r.had.co.nz/Data-structures.html) by Hadley Wickham

<br/>
Understanding data structures is very boring but very **important** :)



---&twocol

## Atomic Vectors

*** =left
    Vector Types

* __Logical__
* __Integer__
* __Double__
* __Character__
* __Complex__
* __Raw__

*** =right
    Creating Vectors

Vectors are created using the command `c()`    

```{r echo=TRUE,eval=T, comment=NA, cache=TRUE,tidy =TRUE}
log <- c(TRUE,FALSE)
inte <- c(1L,2L)
doub <- c(2.3,4.6,7.9)
string <- c("a","b","c")

```
---
## Structural Tests

Try the commands `typeof()` & `str()`
```{r eval =FALSE}
typeof(logi)
str(inte)
```

Other commands to figure out the structure of a vector are `is.double()`,`is.logical()` etc.
```{r echo=TRUE,eval=T, comment=NA, cache=TRUE, tidy =TRUE}
is.double(doub)
is.double(string)
```
---
## Coercion

### Atomic Vectors are **homogenous**

Try creating a few mixed vectors and determine their types

```{r eval = F,cache=TRUE, tidy=T}
mixed_1 <- c(1,T) 
mixed_2 <- c("Sunanda", 25)
```

Homogenous data structures do not allow different types of data to be clumped together.

---
## Subsetting Vectors

Subsetting means extracting certain observations you are interested in and there are various ways to do it

```{r subsetting,echo=TRUE,eval=F, comment=NA, cache=TRUE,tidy=T}
my_vector <- c("blue","green","yellow")

my_vector[2] 

named_vec <- c(color="green",type="light") ##each element can have a name associated to it

names(named_vec) ##an attribute of the vector

named_vec["color"]

```

---
## Conditional Subsetting

```{r echo=TRUE,eval=T, comment=NA, cache=TRUE,tidy=T}
num_vec <- c(1:10,20:35) ##numeric vector

num_vec[num_vec > 15]

num_vec[num_vec < 15 | num_vec >30] #or seperation


```

---
## Lists

### Lists are a type of vector **BUT** can contain any type of vector i.e. heterogenous
```{r echo=TRUE,eval=-3, comment=NA, cache=TRUE,tidy=TRUE}
my_list <- list(log,doub,inte,string)
new_list <- list(my_list,c(1,2,3,4)) ##list in a list
```

Try out these commands
```{r echo=TRUE,eval =T, comment=NA, cache=TRUE,tidy=TRUE}
str(my_list)
```
```{r eval=FALSE}
is.recursive(new_list)
```
---
## Dataframes
<br/>
- Dataframes are a 'special' type of list where each vector is equi-length
- Dataframes are always 2-dimensional (row and columns)
<br/>
<br/>

### Try to create the dataframe in the same way as the list using the `data.frame()` function

### What error do you get?


---
## Getting Help

For getting the R Documentation on any R function simply type `?[function_name]`

For additional help, usually google is the best resource.

A lot of forums like **stackoverflow** have very active R user communities.

---
## Working with a Dataframe

#### Setting the working directory
```{r wd, eval = F, echo=T,tidy =TRUE, cache=T}
## set the working directory
setwd("//filer001.office.nugg.ad/Data/Statistik/RPTN/Learning/R/Lerngruppe")

my_data <- read.table("data_s.csv",sep="\t",header=T)

```

**read.table() Options**
- _file_ :name of the file to be read
- _sep_ :how the data is seperated
- _header_ :whether data has column names of not
.. and many others. Try ?read.table

---

## Subsetting Dataframe

R has an indexing system, which makes it pretty easy to extract information


```{r eval = F, echo=T,tidy =TRUE, cache =T}
##4th row, & 5th column
my_data[4,5]

new_data <- subset(my_data, s4n > 20)

dim(new_data) ##dimensions of the newly created data

```

---middle
## Thats all for this time..


For more information: Advanced R Programming by Hadley Wickham (Chapter: Data Structures)








